---
title: "Reproducible Research: Peer Assessment 1"
author: Joe XU
date: 27/03/2020
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
require(data.table)
require(ggplot2)

activity <- read.csv("./repdata_data_activity/activity.csv")
activity <- as.data.table(activity)
```

## What is mean total number of steps taken per day?

Calculate the total number of steps taken per day
Make a histogram of the total number of steps taken each day
```{r}
# Calculate the total number of steps taken per day
totalSteps <- activity[ , .(totalSteps = sum(steps, na.rm = TRUE)), by = .(date)]

#hist(meanSteps$avgSteps, xlab = "average steps", ylab = "No. of days")
ggplot(totalSteps, aes(totalSteps)) +
    geom_histogram(bins = 50) + 
    labs(x = "Number of Steps", y="Number of Days", 
         title = "Total number of steps taken per day")
```

Mean and Median steps taken per day

```{r}
print(paste("mean steps per day: ", 
            round(mean(totalSteps$totalSteps), digits = 0)))
print(paste("media steps per day: ", 
            round(median(totalSteps$totalSteps), digits = 0)))      
```

## What is the average daily activity pattern?
Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
actInterval <- activity[ , .(meanIntSteps = mean(steps, na.rm = TRUE)), by = .(interval)]
ggplot(na.omit(actInterval), aes(x=interval, y=meanIntSteps)) +
    geom_line(color="blue") +
    labs(x="A Day (5 min Interval)", y="Steps") +
    labs(title="Daily Average Activity Pattern")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
#actInterval[which.max(actInterval$meanIntSteps), meanIntSteps]
actInterval[which.max(actInterval$meanIntSteps), interval]
```

## Imputing missing values
Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
summary(is.na(activity))
```

Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

```{r}
imputeSteps <- function(DF, i) {
  sel <- DF$interval == DF$interval[i]
  imputed <- mean(DF$steps[sel], na.rm = TRUE)
  if (is.nan(imputed)) {
    imputed <- 0
  }
  return(imputed)
}

# apply the function to replace NA values with the mean for that 5 mins interval

for (i in which(is.na(activity$steps))) {
  activity$steps[i] <- imputeSteps(activity, i)
}

summary(is.na(activity))
```

Create a new dataset that is equal to the original dataset but with the missing data filled in. The new dataset is named "activity1.csv"

```{r}
fwrite(x = activity, file = "activity1.csv", quote = FALSE)
```

Make a histogram of the total number of steps taken each day. 

```{r}
# Calculate the total number of steps taken per day
totalSteps <- activity[ , .(totalSteps = sum(steps, na.rm = TRUE)), by = .(date)]

#hist(meanSteps$avgSteps, xlab = "average steps", ylab = "No. of days")
ggplot(totalSteps, aes(totalSteps)) +
    geom_histogram(bins = 50) + 
    labs(x = "Steps", y="Number of Days",
         title = "Total number of steps taken per day")
```

Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
print(paste("mean steps per day: ", 
            round(mean(totalSteps$totalSteps), digits = 0)))
print(paste("media steps per day: ", 
            round(median(totalSteps$totalSteps), digits = 0)))      
```

## Are there differences in activity patterns between weekdays and weekends?


```{r}
activity$type <- ifelse(weekdays(as.Date(activity$date)) %in% c("Saturday", "Sunday"), "weekend", "weekday")
activity$type <- as.factor(activity$type)

actInterval <- activity[ , .(meanIntSteps = mean(steps, na.rm = TRUE)), by = .(interval, type)]

ggplot(actInterval, aes(x=interval,y=meanIntSteps,color=type)) +
    geom_line() +
    facet_grid(scales="free", space="free", type~.) +
    labs(x="A Day (5 min Interval)", y="Steps") +
    labs(title="Daily Average Activity Pattern")
```